Effect of multiple reflection of the sky signal off the HERA element on the measurements of 21cm power spectrum

### 2015-11-02 DishReflectometry/Paper1_v0.tex
Monday update2

\section{Effects of multiple reflections of the sky signal between the parabolic dish and the feed on the measured visibility}
 Plane waves incident on a parabolic dish are focussed at the feed which is kept at the focal plane of the dish. The mismatch between the impedance of free space and the feed and transmission line results in a partial coupling of the sky signal into the feed while the rest is reflected off the feed. The reflected signal illuminates the dish and most of it is reflected back into the space. However, a part of it reflects back and forth several times in between the feed and the vertex of the dish which is shadowed by the feed (dashed blue arrows in Figure \ref{fig:cartoon}). Such reflections generate multiple copies of the incident sky signal of reduced strength at various time/phase and thus produces spurious correlations in the visibilities of interferometric data. In this section we compute the visibility of a two element interferometer accounting for the additional reflections of the sky signal in between the feed and the dish vertex and the corresponding effects on the delay spectrum.

@@

To first order, the design specification of HERA elements requires that any
delayed signal, arriving at the feed after a multipath propagation, should be
at the level of $-60dB$ at a delay of $60ns$ relative to the first incident
signal at the feed \citep{parsons_deboer_memo}. This specification was
approximated based on the power level of the cosmological signal in relation to
foreground signals, which is estimated to be six orders of magnitude fainter
\citep{santos_et_al2005,ali_et_al2008,deoliveira2008,jelic_et_al2008,bernardi_et_al2009,bernardi_et_al2010,ghosh_et_al2011}.
Additionally, the $14m$ HERA baselines set a foreground containing
horizon-limit (the wedge) that, with some buffer, sets a delay specification of
$60ns$
\citep{parsons_et_al2012b,vedantham_et_al2012,nithya_et_al2013,liu_et_al2014a,liu_et_al2014b}.

We carry out reflectometry measurements at the HERA element prototype in Green
Bank, WV (Figure \ref{fig:heradish}) in order to understand the nature of feed
reflections in the dish and characterize its performance. As HERA progresses as
an experiment, it is necessary to build optimized dishes that aim to minimize the
challenges of chromaticity in our quest for the Epoch of Reionization.

Plane waves incident on a parabolic dish
are focussed at the feed with focal height $l$. For a well-designed feed, one
that closely matches the impedance of free space\footnote{The impedance of free space
is $Z_{0} = \mu_{0}c_{0} = \frac{1}{\epsilon_{0}c_{0}} \approx 377\Omega $.},
most of the signal will enter the system while only
a small percentage will be reflected back towards the dish for a secondary
reflection into the feed (dashed blue arrows in Figure \ref{fig:cartoon}). In the
following discussion, we consider a reflection off the feed and the
subsequent reflection off the dish as one reflection.

Quantitatively, if the incident power from the sky signal is $I_{sky}$, the feed
reflection coefficient is $\Gamma_{f}$, and the dish reflection
coefficient is $\Gamma_{d}$, then the net power entering the feed after an
$n^{th}$ reflection off the feed and the dish is:

\begin{equation}\label{eqn:series1}
P_{in} =  P_{sky}(1-\Gamma_{f})[1+ \Gamma_{f}\Gamma_{d} e^{i\phi}+ (\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}]
\end{equation}

where, $\phi = 2l({2\pi \over c})\nu$ is the propagation delay of a lightwave of frequency $\nu$ due to a reflection over a focal distance $l$. It is noted that upon the first incidence, the sky signal is focussed onto the feed from the entire dish and the dish reflection coefficient $\Gamma_{d}$ is $1$ for this first incidence. The subsequent back and forth reflection of the signal in between the feed and the dish, however, occurs from only the part of the dish which is shadowed by the feed. Therefore, in this case, $\Gamma_{d} < 1$.

Simplifying Equation \ref{eqn:series1}:
\begin{eqnarray}\label{eqn:ratio1}
{P_{in} \over P_{sky} } & = & (1-\Gamma_{f})[1+ \Gamma_{f}\Gamma_{d} e^{i\phi}+ (\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}] \nonumber\\
      & = & (1-\Gamma_{f}) \displaystyle\sum\limits_{n=0}^{n} [\Gamma_{f}\Gamma_{d}e^{i \phi}]^{n}\nonumber\\
      & = & (1-\Gamma_{f})+(1-\Gamma_{f}) \displaystyle\sum\limits_{n=1}^{n} [\Gamma_{f}\Gamma_{d}e^{i \phi}]^{n}
\end{eqnarray}

The ratio in Equation \ref{eqn:ratio1} quantifies the amount of power received
with respect to the incident sky power. Realistically, the incident sky power
is not easily quantifiable, but it is a quantity we need to know to accurately
characterize reflections.  
%The right hand side of this equation consists of system parameters which are all chromatic i.e, functions of frequency.

%Reflections cause the sky signal to appear at various delays. We aim to measure
%the relative signal strength at those various delays, referred to as the
%``delay spectrum" hereafter, to estimate levels of reflections in between the
%feed and the dish apex.

%We employ the reciprocity theorem for antennas in our experimental set-up.
%More specifically, our experimental set-up differs from measuring sky signal
%reflections since we send a pulse that is to be transmitted from the feed
%before receiving it again. This results in a reversed situation - most of our
%original signal is transmitted from the feed to be reflected from the dish,
%while a small percentage reflects back down the cable (red arrows in Figure
%\ref{fig:cartoon}). Therefore, since our measurement is carried out in
%transmitting mode while an observation would be carried out in receiving mode,
%we correct our results in order to map them to represent real observations.

In our experimental set-up, instead of using sky signal, we employ
our feed as a transmitter and transmit a pulse. If the initial pulse is a broadband signal,
$P_{tr}$, sent to the feed via a $75ft$ cable by a vector network
analyser (VNA), a delay domain measurement of the system is accomplished by
taking the Fourier transform of the measured complex return loss of the feed. When the signal is incident on
the feed, part of the incident power ($\Gamma_{f}$) is reflected back to the
measuring device (dashed red arrows in Figure \ref{fig:cartoon}) and
$(1-\Gamma_{f})$ is radiated by the feed (solid red arrows in Figure
\ref{fig:cartoon}) which illuminates the dish. While most of the incident signal is reflected back into space by the dish, the reflected signal from the dish vertex returns to
the feed. This incident signal is now
reflected back and forth in between the feed and the dish much like the sky
signal reflection discussed previously.  Hence, if $P_{r}$ is the power
incident back on the feed for the first time then the reflected power, $P_{refl}$,
back into the VNA would be:

\begin{equation}\label{eqn:series2}
P_{refl} =  P_{r}(1-\Gamma_{f})[1+ \Gamma_{f}\Gamma_{d} e^{i\phi}+ (\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}]
\end{equation}

@@

Recall that $P_{r}$ is the initial power that is incident back on the feed, which is just the feed radiated power reflected off the dish:

\begin{equation}
P_{r}= \Gamma_{d}(1-\Gamma_f)e^{i\phi} P_{tr}
\end{equation}

@@

\begin{eqnarray}
P_{ret} & = & \Gamma_{f}P_{tr} \nonumber\\ 
 & + &   \Gamma_{d}(1-\Gamma_f)e^{i\phi} P_{tr}(1-\Gamma_{f}) [1+ \Gamma_{f}\Gamma_{d} e^{i\phi}+  ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}]\nonumber\\
 \end{eqnarray}

@@

  \begin{eqnarray}\label{eqn:ratio2}
 {P_{ret} \over P_{tr} } & = & \Gamma_{f}
  +  \Gamma_{d}(1-\Gamma_f)^{2} [e^{i\phi}+ \Gamma_{f}\Gamma_{d} e^{i2\phi}+  ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{i n\phi}]\nonumber\\
  & = & \Gamma_{f} + { (1-\Gamma_f)^{2}\over \Gamma_{f} } \displaystyle\sum\limits_{n=1}^{n} [\Gamma_{f}\Gamma_{d}e^{i\phi}]^{n}
   \nonumber\\
\end{eqnarray}

@@

The VNA measures the magnitude and phase of the quantity ${P_{ret}\over P_{tr}}$
as a function of frequency as shown in Figure \ref{fig:freq}. In our measurement
set-up, the first reflection occurs at the feed $\Gamma_{f}$, so
$({P_{ret} \over P_{tr} }  - \Gamma_{f}) $ gives an estimate of the delay
spectrum of the sky signal. In delay domain, the relative signal strength at
zero delay represents the factor $\Gamma_{f}$ while the signal strength at any
other delay represents any delayed signal that enters the feed after being
reflected from the feed surroundings.

>>>>

\section{Effects of additional reflections of the sky signal between the parabolic dish and the feed on the measured visibility}
 Plane waves incident on a parabolic dish are focussed at the feed which is kept at the focal plane of the dish. The mismatch between the impedance of free space and the feed and transmission line results in a partial coupling of the sky signal into the feed while the rest is reflected off the feed. The reflected signal illuminates the dish and most of it is reflected back into the space. However, a part of it reflects back and forth several times in between the feed and the vertex of the dish which is shadowed by the feed (dashed blue arrows in Figure \ref{fig:cartoon}). Such reflections generate multiple copies of the incident sky signal of reduced strength at various time/phase and thus produces spurious correlations in the visibilities of interferometric data. In this section we compute the visibility of a two element interferometer accounting for the additional reflections of the sky signal in between the feed and the dish vertex and the corresponding effects on the delay spectrum.

@@

and the measured visibility spectrum, integrated over all sky would be, 
\begin{equation}\label{eqn:series1}
V_{meas}  = I_{sky}A_{f}\Sigma( \nu)
\end{equation}

The design specification of HERA elements requires that any
signal, arriving at the feed at a delay $\Delta \tau$ with respect to the direct incidence, should be
at the level of $-60dB$ at a delay of $60ns$ relative to the first incident
signal at the feed \citep{parsons_deboer_memo}. This specification was
approximated based on the power level of the cosmological signal in relation to
foreground signals, which is estimated to be six orders of magnitude fainter
\citep{santos_et_al2005,ali_et_al2008,deoliveira2008,jelic_et_al2008,bernardi_et_al2009,bernardi_et_al2010,ghosh_et_al2011}. Additionally, the $14m$ HERA baselines set a foreground containing
horizon-limit (the wedge) that, with some buffer, sets a delay specification of
$60ns$
\citep{parsons_et_al2012b,vedantham_et_al2012,nithya_et_al2013,liu_et_al2014a,liu_et_al2014b}. Our goal is to determine the effect of the instrument response $\Sigma(\nu)$ on the delay spectrum measurement. 
\section{Measurements}
We carried out reflectometry measurements at the HERA element prototype in Green
Bank, WV (Figure \ref{fig:heradish}) in order to measure the instrument response of the feed and dish assembly and characterise its performance. As HERA progresses as
an experiment, it is necessary to build optimised dishes that aim to minimise the
challenges of chromaticity in our quest for the Epoch of Reionization.

As given in the previous section, if the incident power from the sky signal is $I_{sky}$, the feed
reflection coefficient is $\Gamma_{f}$, and the dish reflection
coefficient is $\Gamma_{d}$, then the net power entering the feed after
$n^{th}$ reflection off the feed and the dish (i.e, the sky integrated visibility spectrum) is:

\begin{equation}\label{eqn:series1}
I_{meas} =  I_{sky}A_{f}[1+ \Gamma_{f}\Gamma_{d} e^{i\phi}+ (\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}]
\end{equation}

where, $\phi = 2\pi i\nu \Delta \tau$ is the roundtrip propagation delay of a lightwave of frequency $\nu$ due to a reflection over the focal distance $l$. It is noted that upon the first incidence, the sky signal is focussed onto the feed from the entire dish and the dish reflection coefficient $\Gamma_{d}$ is $1$ for this first incidence. The subsequent back and forth reflection of the signal in between the feed and the dish, however, occurs from only the part of the dish which is shadowed by the feed. Therefore, in this case, $\Gamma_{d} < 1$.

Simplifying Equation \ref{eqn:series1}:
\begin{eqnarray}\label{eqn:ratio1}
{I_{in} \over I_{sky} } & = & A_{f}[1+ \Gamma_{f}\Gamma_{d} e^{i\phi}+ (\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}] \nonumber\\
      & = & A_{f} \displaystyle\sum\limits_{n=0}^{n} [\Gamma_{f}\Gamma_{d}e^{i \phi}]^{n}\nonumber\\
      & = & A_{f}+A_{f} \displaystyle\sum\limits_{n=1}^{n} [\Gamma_{f}\Gamma_{d}e^{i \phi}]^{n}
\end{eqnarray}

In our experimental set-up, we use the feed as a transmitter and transmit a pulse. If the initial pulse is a broadband signal,
$I_{tr}$, sent to the feed via a $75ft$ cable by a vector network
analyser (VNA), a delay domain measurement of the system is accomplished by
taking the Fourier transform of the complex return loss of the feed measured by VNA. When the signal is incident on
the feed, part of the incident power ($\Gamma_{f}$) is reflected back to the
measuring device (dashed red arrows in Figure \ref{fig:cartoon}) and
$(1-\Gamma_{f})$ is radiated by the feed (solid red arrows in Figure
\ref{fig:cartoon}) which illuminates the dish. While most of the incident signal is reflected back into space by the dish, the reflected signal from the dish vertex returns to
the feed. This incident signal is now
reflected back and forth in between the feed and the dish much like the sky
signal reflection discussed previously.  Hence, if $I_{r}$ is the power
incident back on the feed for the first time then the reflected power, $I_{refl}$,
back into the VNA would be:

\begin{equation}\label{eqn:series2}
I_{refl} =  I_{r}A_{f}[1+ \Gamma_{f}\Gamma_{d} e^{i\phi}+ (\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}]
\end{equation}

@@

Recall that $I_{r}$ is the initial power that is incident back on the feed, which is just the feed radiated power reflected off the dish:

\begin{equation}
I_{r}= \Gamma_{d}(1-\Gamma_f)e^{i\phi} I_{tr}
\end{equation}

@@

\begin{eqnarray}
I_{ret} & = & \Gamma_{f}I_{tr} \nonumber\\ 
 & + &   \Gamma_{d}(1-\Gamma_f)e^{i\phi} I_{tr}(1-\Gamma_{f}) [1+ \Gamma_{f}\Gamma_{d} e^{i\phi}+  ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}]\nonumber\\
 \end{eqnarray}

@@

  \begin{eqnarray}\label{eqn:ratio2}
 {I_{ret} \over I_{tr} } & = & \Gamma_{f}
  +  \Gamma_{d}(1-\Gamma_f)^{2} [e^{i\phi}+ \Gamma_{f}\Gamma_{d} e^{i2\phi}+  ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{i n\phi}]\nonumber\\
  & = & \Gamma_{f} + { (1-\Gamma_f)^{2}\over \Gamma_{f} } \displaystyle\sum\limits_{n=1}^{n} [\Gamma_{f}\Gamma_{d}e^{i\phi}]^{n}
   \nonumber\\
\end{eqnarray}

@@

The VNA measures the magnitude and phase of the quantity ${I_{ret}\over I_{tr}}$
as a function of frequency as shown in Figure \ref{fig:freq}. In our measurement
set-up, the first reflection occurs at the feed $\Gamma_{f}$, so
$({I_{ret} \over I_{tr} }  - \Gamma_{f}) $ gives an estimate of the delay
spectrum of the sky signal. In delay domain, the relative signal strength at
zero delay represents the factor $\Gamma_{f}$ while the signal strength at any
other delay represents any delayed signal that enters the feed after being
reflected from the feed surroundings.
